---
title: "Final_Project"
author: "Wilson Encarnacion"
format: html
editor: visual
---

```{r}
library(tidyverse)
ev_data <- read_csv("https://data.wa.gov/api/views/f6w7-q2d2/rows.csv?accessType=DOWNLOAD")

```
#Average electric range
```{r}
avg_range <- ev_data %>% 
  summarize(avg_range = mean(`Electric Range`, na.rm = TRUE))
print(avg_range)
```
The average electric range of vehicles in the data is 45.56 miles. This suggests that most EVs in Washington State have a relatively modest range, which could be a result of a mix of BEVs and PHEVs.

To further examine the distribution of electric range, a histogram is used:
#Histogram
```{r}
ggplot(ev_data %>% drop_na(`Electric Range`), aes(x = `Electric Range`)) +
  geom_histogram(binwidth = 10, fill = "blue", alpha = 0.7) +
  theme_minimal() +
  labs(title = "Distribution of Electric Vehicle Ranges", 
       x = "Electric Range (Miles)", 
       y = "Count of Vehicles")

```

The histogram shows the variation in electric range across different EV models. Peaks in certain ranges may indicate popular models with specific battery capacities.

#Two-sample t-test
To test whether BEVs and PHEVs differ significantly in MSRP, a Welch Two-Sample t-test is performed:
```{r}
t_test_result <- t.test(`Base MSRP` ~ `Electric Vehicle Type`, data = ev_data)
print(t_test_result)
```
The p-value < 2.2e-16, meaning the difference in MSRP is highly significant.

BEVs have a lower average MSRP ($640.78), while PHEVs are significantly higher ($1267.24).


#Boxplot
The boxplot compares the distribution of the MSRP
```{r}
ggplot(ev_data %>% drop_na(`Base MSRP`), aes(x = `Electric Vehicle Type`, y = `Base MSRP`, fill = `Electric Vehicle Type`)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "MSRP Comparison Between BEVs and PHEVs") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```

